Follow the Flow of Rain and Paint: \break An Output Sensitive Contour Tree Algorithm

### 2014-06-29 paint-contour-tree.tex
Minor section 2 changes

\begin{proof} \Clm{mono} should help us argue that for distinct $\phi, \phi' \in \Phi$,
for any $t \in [f(\phi),f(\phi')]$, there exists $\phi'' \in \Phi$ such that $f(\phi'') = t$.
So $f(\Phi)$ is an interval. \Clm{reg} should tell us that this interval must be open
with critical value endpoints.
\end{proof}

@@

\begin{definition} \label{def:reeb} The \emph{Reeb Graph} $\reeb(\MM)$ has as vertex set
the set of critical points. The edge $(x,y)$ is present if there exists equivalence class
$\Phi$ such that $f(\Phi) = [f(x),f(y)]$.
\end{definition}

@@

\smallskip
\begin{asparaenum}
%	\item Insert the vertices of $\phi$ in $\MM$.
	\item Take an arbitrary edge $e$ intersecting $\phi$.
	\item Take a point $x$ (resp. $y$) on $e$ at distance $\eps$ above (resp. below) $\phi$.
	\item Let $\phi_x$ be the contour through $x$, and analogously define $\phi_y$. Insert a vertex
	at each intersection point of these contours with edges of the manifold. Triangulate all new faces.
	\item Delete the vertices (and incident edges, faces, etc.) of $\phi$ from the manifold.
\end{asparaenum}
\end{minipage}}

\medskip
This results in a new manifold that is also boundary critical. There are two new disjoint
boundary faces added. The manifold may be disconnected. Abusing notation, we use $\cut(\MM,\Phi)$
for a set $\Phi$ of contours to mean the manifold obtained by cutting along all contours in $\Phi$.

@@

\begin{definition} \label{def:wet} The set of points $y$ such that there is a descending path from $x$ to $y$
is denoted by $\wet(x,\MM)$. A point $z$ is on the \emph{interface} of $\wet(x,\MM)$ if every neighbor
has non-trivial intersection with $\wet(x,\MM)$.
\end{definition}

@@

\begin{proof} If $y \in \wet(x,\MM)$, then any contour containing $y$ is also in $\wet(x,\MM)$.
Consider a $y$ in the interface. We argue that all points in a contour $\phi$ containing $y$
are also in the interface. If such a contour is regular, we can find a contour just above it that
is also wet.

If two contours intersect, the intersection point $y$ is critical. We can argue that the only
way to reach $y$ from a descending path is reach one of the contours and then follow it to get to $y$.
But that would imply the existence of another critical point on this contour, violating the Morse condition.

Let us argue that $y$ is a merge vertex. Consider the descending path from $x$ to $y$. The final edge
that reaches $y$ is obviously all wet, and is an increasing edge from $y$. Consider an opposite increasing
edge. Take two point in these edges that are infinitesimally higher than $y$ and show that the contours
through them as disjoint. So $y$ is a merge.
\end{proof}

We will define a \lift{} operation on the interface contours. Consider such a contour $\phi$ containing
merge vertex $y$. Take any dry increasing edge incident to $y$, and pick the point $z$ on this edge at height
$f(y) + \eps$. Let $\lift(y) = \cont(z)$.

\begin{claim} \label{clm:cut-int} Let $\phi$ be an interface contour. Then $\cut(\MM,\lift(\phi))$
results in two disjoint manifolds, one consisting entirely of dry points, and the other containing
all wet points.
\end{claim}

\begin{proof} Note that $\lift(\phi)$ is a closed, dry curve. But the Jordan Curve Theorem,
$\cut(\MM,\lift(\phi))$ results in two disjoint manifolds. Let $\NN$ be the manifold containing
the point $x$, and $\NN'$ be the other manifold. If there exists a wet point $z$ in $\NN'$,
then there exists a descending path from $x$ to $z$ intersecting $\phi$ (Jordan Curve Theorem again).
That contradicts the fact that $\phi$ is dry.
\end{proof}

@@

\smallskip
\begin{compactenum}
	\item Compute the interface of $\wet(x,\NN)$. 
	\item If the interface is empty, simply output $\NN$. Otherwise, denote the contours by $\phi_1, \phi_2, \ldots, \phi_k$, and set $\phi'_i = \lift(\phi_i)$.
	\item Initialize $\NN_1 = \NN$.
	\item For $i$ from $1$ to $k$: 
	\begin{compactenum}
		\item Construct $\cut(\phi'_i,\NN_1)$ to get the dry manifold $\LL_i$ and wet manifold $\NN_{i+1}$.
		\item Let the new boundary of $\LL_i$ be $B_i$. Invert $\LL_i$ so that $B_i$ is a maximum. Recursively
		call $\rain(B_i,\LL_i)$.
	\end{compactenum}
	\item Output $\NN_{k+1}$.
\end{compactenum}
\end{minipage}}

>>>>

\begin{proof} \Clm{mono} and continuity of L-homotopy immediately imply that for $\phi, \phi' \in \Phi$,
for any $t \in [f(\phi),f(\phi')]$, there exists $\phi'' \in \Phi$ such that $f(\phi'') = t$.
So $f(\Phi)$ is an interval. As equivalence classes are defined over only regular contours, \Clm{reg} 
tells us that this interval must be open with critical value endpoints.
\end{proof}

@@

\begin{definition} \label{def:reeb} The \emph{Reeb Graph} $\reeb(\MM)$ has as a vertex set
the set of critical points. The edge $(x,y)$ is present if there exists equivalence class
$\Phi$ such that $f(\Phi) = (f(x),f(y))$.
\end{definition}

@@

\smallskip
\begin{asparaenum}
%	\item Insert the vertices of $\phi$ in $\MM$.
	\item Take an arbitrary edge $e$ whose interior intersect $\phi$.
	\item Take a point $x$ (resp. $y$) on $e$ at distance $\eps$ above (resp. below) $\phi$.
	\item Let $\phi_x$ be the contour through $x$, and analogously define $\phi_y$. Insert a vertex
	at each intersection point of these contours with edges of the manifold. Triangulate all new faces.
	\item Delete the vertices (and incident edges, faces, etc.) of $\phi$ from the manifold.
\end{asparaenum}
\end{minipage}}

\medskip
This results in a new manifold which is boundary critical if the input manifold was. There are two new disjoint
boundary faces added. The manifold may be disconnected. Abusing notation, we use $\cut(\MM,\Psi)$
for a set $\Psi$ of contours to mean the manifold obtained by cutting along all contours in $\Psi$.

@@

\begin{definition} \label{def:wet} The set of points $y$ such that there is a descending path from $x$ to $y$
is denoted by $\wet(x,\MM)$. A point $z$ is on the \emph{interface} of $\wet(x,\MM)$ if every neighborhood of $z$
has non-trivial intersection with $\wet(x,\MM)$.
\end{definition}

@@

\begin{proof}
 First observe that any point which in the interface must be wet.  Second, for $y \in \wet(x,\MM)$, all the points in any contour containing $y$ are also in $\wet(x,\MM)$, 
 as one can follow the descending path to $y$ and then walk along the contour.

 So let $y$ be a point in the interface.
 For $\eps$ sufficiently small, the points in any $\eps$ neighborhood of $y$ that lie below $y$ must have a descending path from $y$ and hence must be wet.
 Therefore in order for $y$ to have non-trivial intersection with $\wet(x,\MM)$ for any arbitrarily small $\eps$ there must be a dry point which lies 
 strictly above $y$.  In particular, this implies there must be an entire dry contour which lies just above $y$.

 However, since $y$ is in the interface, it is also wet.  Therefore for any contour $\phi$ which contains $y$, for any arbitrarily small $\eps$ there is 
 a wet point which lies above $\phi$ and has a descending path of length at most $\eps$ to $\phi$.
 This implies there is some contour which lies just 
 above $y$ which is wet.  So any contour $\phi$ containing $y$ has two contours which lie just above it, one which is dry, and one which is wet.  
 This can only happen if $\phi$ contains a merge vertex, because otherwise all contours which lie just above would be in the same L-homotopy equivalence class and so by 
 \Clm{mono}, if one is wet, all below it are wet.

 By similar logic one can argue that of the two contours through a given merge vertex at most one can contain interface points other than the merge vertex, 
 and that if one of the contours does contain such interface points, then the entire contour consists of interface points.
\end{proof}

% \begin{proof} If $y \in \wet(x,\MM)$, but not in the interface, then any contour containing $y$ is also in $\wet(x,\MM)$.
% So consider a point $y$ in the interface. We argue that all points in a contour $\phi$ containing $y$
% are also in the interface. If such a contour is regular, we can find a contour just above it that
% is also wet. 
% 
% If two contours intersect, the intersection point $y$ is critical. We can argue that the only
% way to reach $y$ from a descending path is reach one of the contours and then follow it to get to $y$.
% But that would imply the existence of another critical point on this contour, violating the Morse condition.
% 
% Let us argue that $y$ is a merge vertex. Consider the descending path from $x$ to $y$. The final edge
% that reaches $y$ is obviously all wet, and is an increasing edge from $y$. Consider an opposite increasing
% edge. Take two point in these edges that are infinitesimally higher than $y$ and show that the contours
% through them as disjoint. So $y$ is a merge.
% \end{proof}

We will define a \lift{} operation on the interface contours. Consider such a contour $\phi$ containing
a merge vertex $y$. Take any dry increasing edge incident to $y$, and pick the point $z$ on this edge at height
$f(y) + \eps$. Let $\lift(y) = \cont(z)$.

\begin{claim} \label{clm:cut-int} Let $\phi$ be an interface contour. Then $\cut(\MM,\lift(\phi))$
results in two disjoint manifolds, one consisting entirely of dry points.
\end{claim}

\begin{proof} Note that $\lift(\phi)$ is a closed, dry curve. By the Jordan Curve Theorem,
$\cut(\MM,\lift(\phi))$ results in two disjoint manifolds. Let $\NN$ be the manifold containing
the point $x$ that we called $\wet(x,\MM)$ from, and let $\NN'$ be the other manifold. 
If there exists a wet point $z$ in $\NN'$, then there exists a descending path from $x$ to $z$. 
As $x$ is in $\NN$ and $z$ is in $\NN'$, again by the Jordan Curve Theorem this path must intersect $\phi$, 
contradicting the fact that $\phi$ is dry.
\end{proof}

@@

\smallskip
\begin{compactenum}
	\item Compute the interface of $\wet(x,\NN)$. 
	\item If the interface is empty, simply output $\NN$. Otherwise, denote the contours by $\phi_1, \phi_2, \ldots, \phi_k$, and set $\phi'_i = \lift(\phi_i)$.
	\item Initialize $\NN_1 = \NN$.
	\item For $i$ from $1$ to $k$: 
	\begin{compactenum}
		\item Call $\cut(\phi'_i,\NN_1)$ to get the dry manifold $\LL_i$ and wet manifold $\NN_{i+1}$.
		\item Let the new boundary of $\LL_i$ be $B_i$. Invert $\LL_i$ so that $B_i$ is a maximum. Recursively
		call $\rain(B_i,\LL_i)$.
	\end{compactenum}
	\item Output $\NN_{k+1}$.
\end{compactenum}
\end{minipage}}
